### 2016-11-07 papers/hrr-paper/bibliography.tex

\begin{thebibliography}{9}
	\bibitem{vsa1}
		"Vector Symbolic Architectures: A New Building Material for Artificial General Intelligence", Simon D. Levy, Ross W Gayler, Frontiers in Artificial Intelligence and Applications, January 2008

		\bibitem{geometric-algebra}
		"Distributed Representations Based on Geometric Algebra: The Continuous Model", Agnieszka Patyk-Lonska, Marek Czachor, Diederik Aerts, Informatica 35(2011), pg. 407 - 417

	\bibitem{pattern-recognition}
	"Pattern Recognition with Vector Symbolic Architectures", Kleyko Denis, Lulea tekniska universitet, Master thesis 2016.

	\bibitem{holographic-graph-nouron}
	"Holographic Graph Neuron: a Bio-Inspired Architecture forPattern Processing", DenisKleyko et. al, IEEE Transactions on Neural Networks and Learning Systems, March 2016
\end{thebibliography}